---
title: "Ventotene_MR_MigStrategies"
output: word_document
author: "Silke Bauer"
---
Document created and analyses run on `r format(Sys.Date(), "%d %B %Y")`
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "c:/A_Dropbox/Dropbox/MigMalaria_NF/subprojects/Ventotene/Stats/MR_MigStrategy/")

getwd()
```

## General

Preliminary analyses, descriptive statistics etc. 


```{r data_handling}

library(lme4)
library(arm)
#setwd("g:/Dropbox/MigMalaria_NF/subprojects/ventotene/Stats/")
setwd("c:/A_Dropbox/Dropbox/MigMalaria_NF/subprojects/Ventotene/Stats/MR_MigStrategy/")
s.test <- data.frame(read.csv("MR_Parasites_2016_2017_data_21062019.csv", h=T, sep=",",fill = T))
s.test <- s.test[1:353,1:38]
s <- s.test
#str(s)

s$mig <- factor(s$mig, levels=c("1","2"), labels=c("short-dist","long-dist"))
s$sex <- factor(s$sex, levels=c("1","2"), labels=c("male","female"))
s$infection <- factor(s$infection, levels=c("0","1"), labels=c("non-inf","infected"))
s$parasitemia <- factor(s$parasitemia, levels=c("0","1","2"), labels=c("no","low","high"))
s$spec <- factor(s$spec, levels=c("1","2","3","4"), labels=c("EriRub","PhoOch","PhoPho","SaxRub"))
s$spec2 <- s$spec

# summary(s)

# The following contrast matrix introduces groups for comparison, with 
# 1. group compares short and long-distance migrants, 
# 2. group compares species within short-distance migrants  
# 3. group compares species within long-distance migrants. 

mat <- model.matrix(~s$spec2, data = s); # mat
contrasts_spec <- matrix(data=c( 0.5, 0.5, 0,
                                 0.5,-0.5, 0,
                                -0.5,   0, 0.5,
                                -0.5,   0,-0.5), nrow=4, ncol=3, byrow=T)
contrasts(s$spec2) <- contrasts_spec


# which data do we use in the following?
s.ex <- s[s$fat > 0,]
d <- s.ex


# Stichprobengr?ssen pro Art & Sex & Infection
table(d$spec, d$sex)
table(d$spec, d$infection)

# Tables give sample sizes for BMR, PMR and all other metabolic-rate-derives output measures
# False heisst "nicht-NA", also g?ltige Werte
table(d$spec, is.na(d$BMR))
table(d$spec, is.na(d$PMR30))
table(d$spec, is.na(d$FAS30))
table(d$spec, is.na(d$Time_exhaust))
table(d$spec, is.na(d$hemo))


```

## Descrptive statistics for subgroups

This part provides means (and standard deviations or other descriptive measures) for body masses, BMR, PMR, FAS, AAS, hemo & time-to-exhaustion for subgroups:
- for short- and longdistance migrants
- for both short-distance migrants
- for both long-distance migrants


```{r descriptives,  echo=TRUE}

# Body masses original
v <- aggregate(cbind(mass_orig_BMR, mass_orig_PMR) ~ sex + spec, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v
v <- aggregate(cbind(mass_orig_BMR, mass_orig_PMR) ~ sex + mig, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v

# BMR and mass-specific BMR
v <- aggregate(cbind(BMR, origMassSpec.BMR) ~ sex + spec, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v
v <- aggregate(cbind(BMR, origMassSpec.BMR) ~ sex + mig, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v

# PMR and mass-specific PMR
v <- aggregate(cbind(PMR30, origMassSpec.PMR) ~ sex + spec, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v
v <- aggregate(cbind(PMR30, origMassSpec.PMR) ~ sex + mig, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v

# FAS and AAS
v <- aggregate(cbind(FAS30, AAS30) ~ sex + spec, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v
v <- aggregate(cbind(FAS30, AAS30) ~ sex + mig, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v

# Time to exhaustion and hemo
v <- aggregate(cbind(Time_exhaust, hemo) ~ sex + spec, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v
v <- aggregate(cbind(Time_exhaust, hemo) ~ sex + mig, data= d, FUN= function(x) c(mean = mean(x), stdev= sd(x))); v



```




## Models for body mass

We calculate the following test:
-	Test for differences in body measures (BMR, PMR) between SD, LD, species
Body mass BMR (Body mass PMR) ~species + sex
Oder besser: gleich mit deiner Kontrastrechnung
Body mass BMR (Body mass PMR) ~migration + sex
???	f?r mass_orig_BMR, massP3_BMR, massTars_BMR
wenn Masse zwischen Arten gleich: -> im weiteren ,normale' MR-Daten und K?rpergr?ssen-skalierte Body Masse -Daten nutzen


```{r bodymass,  echo=FALSE}

# Body mass BMR
lm1.BM <- lm(mass_orig_BMR ~ spec2 + sex, data=d); summary(lm1.BM) 
lm1.BM <- lm(mass_orig_BMR ~ spec2 * sex, data=d); summary(lm1.BM) 

# Body mass PMR 
lm1.BM <- lm(mass_orig_PMR ~ spec2 + sex, data=d); summary(lm1.BM) 
lm1.BM <- lm(mass_orig_PMR ~ spec2 * sex, data=d); summary(lm1.BM) 


# Kruskal-Wallis for fat and muscle
lm1.Fat <- lm(rank(fat) ~ 1 + spec2 + sex, data=d); summary(lm1.Fat) 
lm1.Fat <- lm(rank(fat) ~ 1 + spec2 + infection, data=d); summary(lm1.Fat) 
lm1.Fat <- lm(rank(fat) ~ 1 + spec2 + sex + infection, data=d); summary(lm1.Fat) 
kruskal.test(d$fat ~ d$spec2)

lm1.Muscle <- lm(rank(muscle) ~ 1 + spec2 + sex, data=d); summary(lm1.Muscle) 
lm1.Muscle <- lm(rank(muscle) ~ 1 + spec2 + infection, data=d); summary(lm1.Muscle) 
lm1.Muscle <- lm(rank(muscle) ~ 1 + spec2 + sex + infection, data=d); summary(lm1.Muscle) 
kruskal.test(d$muscle ~ d$spec2)


```


wenn Massen zwischen Arten unterschiedlich: -> Masse-spezifische MR-Daten nutzen (O2 Verbrauch/ g K?rpermasse).
-	Test auf Unterschiede im Hemoglobin zwischen den Strategien/Arten
Hemoglobin ~species + sex
-	Test auf Unterschiede im FatScore und Muscle Score zwischen den Strategien/Arten
Fat~species + sex
Muscle~species + sex


```{r Models_HemoFatMuscle}
 
lm1.Hemo <- lm(hemo ~ spec2 + sex, data=d); summary(lm1.Hemo) 
lm1.Hemo <- lm(hemo ~ spec2 + sex + infection, data=d); summary(lm1.Hemo) 


lm1.Fat <- lm(fat ~ spec2 + sex, data=d); summary(lm1.Fat) 
lm1.Fat <- lm(fat ~ spec2 + sex + infection, data=d); summary(lm1.Fat) 


lm1.Muscle <- lm(muscle ~ spec2 + sex, data=d); summary(lm1.Muscle) 
lm1.Muscle <- lm(muscle ~ spec2 + sex + infection, data=d); summary(lm1.Muscle) 


```

## Plots for BMR as dependent on species, mass, sex


```{r Plots_BMR,  echo=FALSE}

par(mfcol=c(2,2), pch=16)
ymin <- min(log10(d$BMR), na.rm=T); ymax <- max(log10(d$BMR), na.rm=T); 
xmin <- min(d$mass_orig_BMR, na.rm=T); xmax <- max(d$mass_orig_BMR, na.rm=T); 
cols <- c("red","orange","lightblue4","green")


plot(log10(BMR) ~ mass_orig_BMR, data=d[d$spec=="EriRub",], ylim=c(ymin,ymax), xlim=c(xmin,xmax), col=cols[1], main="EriRub")
plot(log10(BMR) ~ mass_orig_BMR, data=d[d$spec=="PhoOch",], ylim=c(ymin,ymax), xlim=c(xmin,xmax), col=cols[2], main="PhoOch")
plot(log10(BMR) ~ mass_orig_BMR, data=d[d$spec=="PhoPho",], ylim=c(ymin,ymax), xlim=c(xmin,xmax), col=cols[3], main="PhoPho")
plot(log10(BMR) ~ mass_orig_BMR, data=d[d$spec=="SaxRub",], ylim=c(ymin,ymax), xlim=c(xmin,xmax), col=cols[4], main="SaxRub")

```


# Models for Metabolic Rates (BMR, PMR, AAS, FAS)
## Models for BMR as dependent on mig strategy, sex, hemo, etc. 
- lgMR~migration+lg Body mass
- lgMR~migration+sex+lg Body mass
- lgMR~migration+hemoglobin+lg Body mass
- lgMR~migration+sex+hemoglobin+lg Body mass

- lgMR~migration+lg Body mass+infection
- lgMR~migration+lg Body mass+infection+hemo


```{r Models_BMR}

# BMR as dependent on species and migration strategy

lm2.BMR <- lm(log10(BMR)     ~ spec2 + log10(mass_orig_BMR),              data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(BMR)     ~ spec2 + log10(mass_orig_BMR) + sex,        data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(BMR)     ~ spec2 + log10(mass_orig_BMR) + hemo,       data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(BMR)     ~ spec2 + log10(mass_orig_BMR) + fat,        data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(BMR)     ~ spec2 + log10(mass_orig_BMR) + muscle,     data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(BMR)     ~ spec2 + log10(mass_orig_BMR) + infection,  data=d); summary(lm2.BMR) 


# Mass-spec BMR as dependent on spec2 and other vairables

lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 ,                                   data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + fat,                              data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + muscle,                           data=d); summary(lm2.BMR) 

lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + sex,                              data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + sex + infection,                  data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + hemo,                             data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + sex + hemo,                       data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + hemo + infection,                 data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + infection,                        data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + sex + hemo + infection,           data=d); summary(lm2.BMR) 


# NEU NEU NEU mit body mass!!!

lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR),           data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + sex,     data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + sex + infection, data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + hemo,   data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + sex + hemo, data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + hemo + infection, data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + infection, data=d); summary(lm2.BMR) 
lm2.BMR <- lm(log10(origMassSpec.BMR) ~ spec2 + log10(mass_orig_BMR) + sex + hemo + infection, data=d); summary(lm2.BMR) 


```



## Models for PMR as dependent on species, mass, sex


```{r Models_PMR}

# PMR as dependent on species and migration strategy

lm1.PMR <- lm(log10(PMR30)            ~ spec2 + log10(mass_orig_PMR), data=d); summary(lm1.PMR) 

lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2,                          data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + fat,                    data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + muscle,                 data=d); summary(lm2.PMR) 

lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + sex,                    data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + hemo,                   data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + infection,              data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + sex + infection,        data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + sex + hemo,             data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + infection + hemo,       data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + sex + infection + hemo, data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + infection * hemo,       data=d); summary(lm2.PMR) 

# NEU NEU NEU NEU mit body mass!!!

lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR),          data=d); summary(lm2.PMR)
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + sex,    data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + hemo,   data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + infection,   data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + sex + infection, data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + sex + hemo,      data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + infection + hemo,data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + sex + infection + hemo, data=d); summary(lm2.PMR) 
lm2.PMR <- lm(log10(origMassSpec.PMR) ~ spec2 + log10(mass_orig_PMR) + infection * hemo, data=d); summary(lm2.PMR) 


```




## Models for AAS, FAS, time-to-exhaustion & hemoglobin as dependent on species, mass, sex


```{r Models_Other}

# Aerobic Scope as dependent on species and migration strategy

lm1.AS <- lm(AAS30 ~ spec2 + log10(mass_orig_BMR),                    data=d); summary(lm1.AS) 
lm1.AS <- lm(AAS30 ~ spec2 + log10(mass_orig_BMR) + sex,              data=d); summary(lm1.AS) 
lm1.AS <- lm(AAS30 ~ spec2 + log10(mass_orig_BMR) + hemo,             data=d); summary(lm1.AS) 
lm1.AS <- lm(AAS30 ~ spec2 + log10(mass_orig_BMR) + hemo + sex,       data=d); summary(lm1.AS) 
lm1.AS <- lm(AAS30 ~ spec2 + log10(mass_orig_BMR) + hemo + infection, data=d); summary(lm1.AS) 

lm1.AS <- lm(FAS30 ~ spec2 + log10(mass_orig_BMR),                    data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + log10(mass_orig_BMR) + sex,              data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + log10(mass_orig_BMR) + hemo,             data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + log10(mass_orig_BMR) + hemo + sex,       data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + log10(mass_orig_BMR) + hemo + infection, data=d); summary(lm1.AS) 

lm1.AS <- lm(FAS30 ~ spec2 + sex,                    data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + hemo,                   data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + infection,              data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + hemo + infection,       data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + sex + infection,        data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + sex + hemo,             data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + sex + hemo + infection, data=d); summary(lm1.AS) 
lm1.AS <- lm(FAS30 ~ spec2 + sex + hemo * infection, data=d); summary(lm1.AS) 


lm5.AS <- lm(Time_exhaust ~ spec2,                          data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + hemo,                   data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + sex,                    data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + infection,              data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + hemo + sex,             data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + hemo + infection,       data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + sex + infection,        data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + infection + sex + hemo, data=d); summary(lm5.AS) 
lm5.AS <- lm(Time_exhaust ~ spec2 + sex + hemo * infection, data=d); summary(lm5.AS) 


```

